---
title: Distribution of Public Service of Canada Employees by Designated Group and Age Range
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
options(warn = -1)
```

## checking if our r is working
```{r}
print("Hello R world")
```

## Analysis of Table 6

### we are loading required libraries
Loading required libraries 
```{r}
#install.packages(c("readxl", "dplyr", "ggplot2", "tidyr"))
library(readxl)
library(janitor)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### loading the data and cleaning the names
we are going to load the data for table 1 and display the first few rows, just to ensure that our data is loaded successfully

we also cleaned the data to use numbers only, excluding the percentages

```{r}
library(readxl)
tab06_eng <- read_excel("~/Documents/assignments/keira/cleaned/tab06-eng.xls", skip = 4, n_max = 7)
Sys.setlocale(category = "LC_CTYPE", locale = "en_US.UTF-8")
tab06_eng <- clean_names(tab06_eng)
print(colnames(tab06_eng))
selected_colnames <- c("age_range", "all_employees", "women", "members_of_a_visible_minority_group","persons_with_disabilities" , "aboriginal_peoples")
subset_data <- tab06_eng[, selected_colnames]
subset_data <- subset_data[complete.cases(tab06_eng$age_range), ]
head(subset_data)
```


### visualization of the data


1. Drawing a bar graph showing the different  distributions of employees against their age ranges

i) Converted the data to numerical data
ii) Created a bar graph

```{r}
subset_data <- subset_data %>%
  mutate_at(vars(all_employees, women, members_of_a_visible_minority_group, persons_with_disabilities, aboriginal_peoples),
            as.numeric)

subset_data_long <- subset_data %>%
  gather(key = "category", value = "value", -age_range)

ggplot(subset_data_long, aes(x = age_range, y = value, fill = category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(title = "Population Distribution by Age Range",
       y = "Population", x = "Age Range") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0, 16000), breaks = seq(0, 16000, by = 1000))
```

from the bar graph above you can see:

- Aboriginal groups are average across the age ranges
- women display a high number across the different ranges



1. distribution of age across populations

```{r}
ggplot(subset_data, aes(x = age_range, y = all_employees, fill = age_range)) +
  geom_bar(stat = "identity") +
  labs(title = "Age Range against total population",
       x = "Age Range",
       y = "Number of All Employees") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

from the bar graph above you can deduce:

- A slighter  working class is under the age of 35
- A set of high number of employees of the age of 50 is represented

2. Scatter plot for distribution of women across different age ranges

```{r}
ggplot(subset_data, aes(x = age_range, y = women)) +
  geom_point() +
  labs(title = "Scatter Plot Age Ranges vs Women",
       x = "Age Range",
       y = "Women") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


from the scatter plot above we can deduce:

- women under the age of 35 is highly represented
- there is a few representation of women between the age of 35-39



3. summary of how the employees are spread out

```{r}
summary_data <- subset_data %>%
  summarise(
    all_employees = sum(all_employees),
    women = sum(women),
    members_of_a_visible_minority_group = sum(members_of_a_visible_minority_group),
    persons_with_disabilities = sum(persons_with_disabilities),
    aboriginal_peoples = sum(aboriginal_peoples)
  )
summary_data_long <- gather(summary_data, key = "category", value = "value")
ggplot(summary_data_long, aes(x = "", y = value, fill = category)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y") +
  labs(title = "Pie Chart of Population Distribution",
       fill = "Category") +
  theme_minimal() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())
```

from the pie chart above we can deduce:

- women are the second most employed category
- Aborginal people and person with disabilities have a few representation in the job industry
